---
title: "Joining Tables"
format: html
editor: visual
---

The exercises in this document follow along closely with the [Joining tables](https://uclouvain-cbio.github.io/WSBIM1207/sec-join.html) chapter of [Introduction to Bioinformatics](https://uclouvain-cbio.github.io/WSBIM1207/index.html), under the creative commons license, [CC-BY 4.0](https://creativecommons.org/licenses/by/4.0/). For more information see:

> Laurent Gatto, Kevin Missault, Manon Martin & Axelle Loriot. (2021, September 24). UCLouvain-CBIO/WSBIM1207: Introduction to bioinformatics (Version v2.0.0). Zenodo. DOI: [10.5281/zenodo.5532658](https://doi.org/10.5281/zenodo.5532658).

<!-- see also https://github.com/UCLouvain-CBIO/WSBIM1207/blob/47465271c0ffc4eac5dd476779a1e5a759a6922e/50-joining-tables.Rmd for the exact version referred to here -->

::: callout-tip
## Key Points

-   Join tables of data in R
:::

```{r}
#| include: false
if(FALSE)
{
    # use this code to install the package we're using for this lesson
    BiocManager::install("UCLouvain-CBIO/rWSBIM1207")
}
library("rWSBIM1207")
data(jdf)

library(dplyr)
```

#### Challenge 1

Using the `full_join` function, join tables `jdf4` and `jdf5`. What has happened for observations `P26039` and `P02468`?

<!-- Challenge 1 answer begin -->

```{r}
# Insert code here
```

<!-- Challenge 1 answer end -->

#### Challenge 2

Join tables `jdf4` and `jdf5`, keeping only observations in `jdf4`.

<!-- Challenge 2 answer begin -->

```{r}
# Insert code here
```

<!-- Challenge 2 answer end -->

#### Challenge 3

Join tables `jdf4` and `jdf5`, keeping only observations in `jdf5`.

<!-- Challenge 3 answer begin -->

```{r}
# Insert code here
```

<!-- Challenge 3 answer end -->

#### Challenge 4

Join tables `jdf4` and `jdf5`, keeping observations observed in both tables.

<!-- Challenge 4 answer begin -->

```{r}
# Insert code here
```

<!-- Challenge 4 answer end -->

#### Challenge 5

Given the inner join,

```{r}
inner_join(jdf6, jdf7)
```

interpret the result, where both tables have duplicated keys.

<!-- Challenge 5 answer begin -->

*insert answer here*

<!-- Challenge 5 answer end -->
